On the Synchronization Power of the OpenFlow Data Plane

### 2015-05-22 inband-sync.tex
committed up to Sec 4

\section{Discussion of Use Cases}\label{sec:apps}

Let us discuss some use cases of our transactional interface.

\subsection{CAS and Consensus}
\petr{ some motivation of why CAS is important/useful is missing}
Given our memory interface and using the bundle concept, we can implement
a powerful \emph{Compare-And-Set} (CAS) operation which takes two integer parameters
$\textit{old}$ and $\textit{new}$, reads the register and, if the
value of it is $\textit{old}$, replaces it with $\textit{new}$ and
returns $\top$; otherwise, $\bot$ is returned and
we say that the CAS operation \emph{fails} (which causes an abort of
the invoking transaction). As can be seen Algorithm \ref{alg:cas}, CAS is implemented by two commands. These commands should be sent as part of the same bundle

@@

    		\State $cmd1\gets \textsc{\compare}(\addr, old) $
    		\State $cmd2\gets \textsc{\memwrite}(\addr, new) $
			\Return $cmd1,cmd2$
    \end{algorithmic}
\end{algorithm}

It is then straightforward to implement consensus among the controllers
using just a single register that can be accessed with an atomic CAS
operation~\cite{Her91}. In our transactional setting, the controllers
can use CAS to replace the default value of such a register
with their consensus input values (e.g., the identifiers of the
policies that should be applied first). If the corresponding
singleton transaction commits, the controller considers itself
the \emph{winner} and decides on its input, if
it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
winner to decide.

>>>>

\section{Use Cases}\label{sec:apps}

We now discuss several applications of our transactional interface.

\subsection{Compare-and-Set}
The pathological scenario described in the example in
Figure~\ref{fig:CAS-example} can be easily resolved if we put
policy update in a transactional $\execatomic}$ construct together
 with a \emph{compare-and-set} (CAS) operation on the policy identifier.
A $\cas(\addr,\textit{old},\textit{new})$ operation on a memory location $\addr$ 
checks if the content of $\addr$ is $\textit{old}$ and if so, 
replaces it with $\textit{new}$ and returns $\true$ (the CAS
\emh{succeeds}), otherwise it returns $\true$ (the CAS
\emh{fails} (which causes an abort of
the invoking transaction). 
It is straightforward to implement the CAS operation by executing
$\textit{\compare}(\addr, old)$ followed by $\textit{\memwrite}(\addr,
new)$ withing the same bundle.

%[[PK said it in the text
\hide{
As can be seen Algorithm \ref{alg:cas}, CAS is implemented by two commands. These commands should be sent as part of the same bundle

@@

    		\State $cmd1\gets \textsc{\compare}(\addr, old) $
    		\State $cmd2\gets \textsc{\memwrite}(\addr, new) $
			\Return $cmd1,cmd2$
    \end{algorithmic}
\end{algorithm}
%]]

CAS can be used to implement a wide range powerful abstractions
shared by the controllers, such as queues, sets, hash tables,
etc.~\cite{Her91} in the \emph{wait-free} manner, i.e., tolerating
asynchrony and failure of any number of controllers.      
In our transactional setting, the controllers
can use CAS to replace the default value of such a register
with their consensus input values (e.g., the identifiers of the
policies that should be applied first). If the corresponding
singleton transaction commits, the controller considers itself
the \emph{winner} and decides on its input, if
it aborts (the controller fails in the CAS) it remains to read the register to get the input of the
winner to decide.
